---
title: "RETORT Rapport"
author: "Hartwig Medical Foundation"
date: "Generated on: `r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    theme: united
params:
  data_file: ""
  iso_fail_file: ""
  prep_fail_file: ""
---
<style type="text/css">
   .main-container {max-width: 80%;}
   .row {display: flex;}
   .column {flex: 50%;}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(dplyr) # Functionality (%>%, filter, select, etc)
library(tidyr) # Table/data frame manipulation
library(ggplot2) # Plotting
library(plotly) # Interactive plots
library(DT) # Interactive tables
library(zoo) # For moving average/median calculations
```

```{r functions, include=FALSE}
calculate_moving_median <- function(dataset, y, coloring) {
  moving_median <- data.frame(matrix(ncol=4, nrow=0))
  sets <- dataset %>% group_split(!!as.name(coloring))
  for (fct_set in sets) {
    fct_set <- fct_set %>% group_by(date, !!as.name(coloring)) %>% summarise(med = median(!!as.name(y))) %>% ungroup()
    mm7_set = as.data.frame(fct_set %>% rollmedian(med, k=7, fill=NA))
    fct_set <- fct_set %>% mutate(mm7 = as.numeric(mm7_set$med))
    moving_median <- rbind(moving_median, fct_set)
  }
  colnames(moving_median) <- c("date", coloring, "med", "mm7")
  return(moving_median)
}

basic_plot <- function(dataset, y, lab_y, lim_y, title, limits_x, coloring) {
  plt <- ggplot(dataset, aes(date, !!as.name(y), color=!!as.name(coloring))) +
    geom_point() +
    ylim(0, lim_y) +
    labs(y=lab_y, x="Date/Time", color=coloring, title=title) +
    scale_x_date(date_labels="%m-%Y") +
    xlim(as.Date(limits_x, format="%Y-%m-%d")) +
    theme_minimal() +
    theme(legend.position="right", axis.text.x=element_text(size=6)) +
    stat_smooth(method="lm", se=FALSE, size=0.5) +
    geom_line(data=calculate_moving_median(dataset, y, coloring), aes(date, mm7, color=!!as.name(coloring)))
  return(plt)
}

basic_table <- function(dataset) {
tbl <- datatable(dataset, selection="multiple", escape=FALSE, rownames=FALSE,
                 options=list(dom='t', pageLength=50, lengthChange=FALSE,
                              columnDefs=list(list(className='dt-center', targets=0:1))))
return(tbl)
}

coefficients_annotation <- function(dataset, x, y, coloring) {
  lm_formula <- paste(y, "~", x)
  annotation <- data.frame(matrix(ncol=2, nrow=0))
  dataset <- dataset[,c(x,y,coloring)]
  for (value in sort(unique(dataset[,coloring]))) {
    reg_data <- dataset %>% filter(!!as.name(coloring) == value)
    regression <- lm(lm_formula, data = reg_data)
    regression <- round(regression$coefficients[-1], 3)
    annotation <- rbind(annotation, c(value, regression))
  }
  colnames(annotation) <- c(coloring, "regression")
  return(annotation)
}
```

```{r load_data, include=FALSE}
# QC Info on Sample level
sample_qc_info <- read.table(params$data_file, sep = '\t', header = TRUE) %>%
                  filter(grepl('^A|L', sequencer))

sample_qc_info$sequencer <- as.factor(sample_qc_info$sequencer)
sample_qc_info["type"][sample_qc_info["type"] == ''] <- "NA"
sample_qc_info$type <- as.factor(sample_qc_info$type)
sample_qc_info$date <- as.Date(sample_qc_info$date, format = '%Y-%m-%dT%H:%M:%S')
sample_qc_info$flowcell_yield <- sample_qc_info$flowcell_yield / 1000000000
sample_qc_info$sample_yield <- sample_qc_info$sample_yield / 1000000000
sample_qc_info$undetermined_percentage <- sample_qc_info$undetermined_percentage / 100

diagnostic_sample_qc_info <- sample_qc_info %>% filter(context == "DIAGNOSTIC", ini == "Somatic.ini")
nondiagnostic_sample_qc_info <- sample_qc_info %>% filter(!(context == "DIAGNOSTIC"), !(ini == "Somatic.ini"))

# QC info on Flowcell level
flowcell_qc_info <- sample_qc_info[!duplicated(sample_qc_info$flowcell),]
diagnostic_flowcell_qc_info <- diagnostic_sample_qc_info[!duplicated(diagnostic_sample_qc_info$flowcell),]
nondiagnostic_flowcell_qc_info <- nondiagnostic_sample_qc_info[!duplicated(nondiagnostic_sample_qc_info$flowcell),]

# Run QC Info
run_qc_info <- as.data.frame(read.csv("/data/tmp/run_qc.csv", sep=",", header=T, na.strings=c("", "NA", "N/A")))
run_qc_info$date <- as.Date(run_qc_info$date, format = '%Y-%m-%dT%H:%M:%S')
run_qc_info$type <- as.factor(run_qc_info$type)
run_qc_info$bqr <- run_qc_info$bqr * 100

run_qc_max_bqr_range <- ceiling(run_qc_info$bqr) + 0.5

# Lab Failures QC Info
iso_fail_info <- read.table(params$iso_fail_file, sep = '\t', header = TRUE)
if (nrow(iso_fail_info) >= 1) {
  iso_fail_info$date <- as.Date(iso_fail_info$date, format = '%Y-%m-%d')
}
prep_fail_info <- read.table(params$prep_fail_file, sep = '\t', header = TRUE)
if (nrow(prep_fail_info) >= 1) {
  prep_fail_info$date <- as.Date(prep_fail_info$date, format = '%Y-%m-%d')
}

# GIAB QC Info
giab_info <- as.data.frame(t(read.csv("/data/tmp/HMF-FOR-096.csv", sep=",", header=F, na.strings=c("", "NA", "N/A"))))
giab_info <- giab_info[,c(2,4,9,12,20,25)]
giab_info <- giab_info[-1,]
colnames(giab_info) <- c("Machine", "SetName", "Q30", "Mapped", "Duplicates", "Excluded")

dates <- as.Date(substr(giab_info[,"SetName"], 1, 6), format='%y%m%d')
giab_info[,"date"] <- dates
giab_info[,"Mapped"] <- as.numeric(gsub("%","",giab_info[,"Mapped"]))
giab_info[,"Duplicates"] <- as.numeric(gsub("%","",giab_info[,"Duplicates"]))
giab_info[,"Excluded"] <- as.numeric(gsub("%","",giab_info[,"Excluded"]))
giab_info[,"Q30"] <- as.numeric(substr(giab_info[,"Q30"], 1, 4))
giab_info$Q30[is.na(giab_info$Q30)] <- 0

giab_info <- giab_info %>% filter(!(is.na(Machine)))
# Date ranges for plotting
qc_date_range <- c(min(flowcell_qc_info$date, na.rm=T), max(flowcell_qc_info$date, na.rm=T))
giab_date_range <- c(min(giab_info$date, na.rm=T), max(giab_info$date, na.rm=T))
```


# Flowcell
### Q30
<div class = "row">
<div class = "column">
```{r diagnostic_flowcell_q30, echo=FALSE}
diag_q30_plot <- basic_plot(
                   diagnostic_flowcell_qc_info,
                   "flowcell_q30",
                   "% Q30",
                   100,
                   "Diagnostic",
                   qc_date_range,
                   "sequencer"
                 ) +
                 geom_hline(yintercept=85, linetype="dashed")
ggplotly(diag_q30_plot)

basic_table(coefficients_annotation(diagnostic_flowcell_qc_info, "date", "flowcell_q30", "sequencer"))
```
</div>
<div class = "column">
```{r nondiagnostic_flowcell_q30, echo=FALSE}
nondiagnostic_q30_plot <- basic_plot(
                flowcell_qc_info,
                "flowcell_q30",
                "% Q30",
                100,
                "Non-Diagnostic",
                qc_date_range,
                "sequencer"
            ) +
            geom_hline(yintercept=85, linetype="dashed")
ggplotly(nondiagnostic_q30_plot)

basic_table(coefficients_annotation(flowcell_qc_info, "date", "flowcell_q30", "sequencer"))
```
</div>
</div>

### Yield
<div class = "row">
<div class = "column">
```{r diagnostic_flowcell_yield, echo=FALSE}
yield_plot <- basic_plot(
                diagnostic_flowcell_qc_info,
                "flowcell_yield",
                "GBase Yield",
                NA,
                "Diagnostic",
                qc_date_range,
                "sequencer"
              )
ggplotly(yield_plot)

basic_table(coefficients_annotation(diagnostic_flowcell_qc_info, "date", "flowcell_yield", "sequencer"))
```
</div>
<div class = "column">
```{r nondiagnostic_flowcell_yield, echo=FALSE}
yield_plot <- basic_plot(
                flowcell_qc_info,
                "flowcell_yield",
                "GBase Yield",
                NA,
                "Non-Diagnostic",
                qc_date_range,
                "sequencer"
              )
ggplotly(yield_plot)

basic_table(coefficients_annotation(flowcell_qc_info, "date", "flowcell_yield", "sequencer"))
```
</div>
</div>

### % Undetermined
<div class = "row">
<div class = "column">
```{r diagnostic_flowcell_undetermined, echo=FALSE}
undetermined_plot <- basic_plot(
                       diagnostic_flowcell_qc_info,
                       "undetermined_percentage",
                       "% Undetermined",
                       100,
                       "Diagnostic",
                       qc_date_range,
                       "sequencer"
                     ) +
                     geom_hline(yintercept=8, linetype="dashed")
ggplotly(undetermined_plot)

basic_table(coefficients_annotation(diagnostic_flowcell_qc_info, "date", "undetermined_percentage", "sequencer"))
```
</div>
<div class = "column">
```{r nondiagnostic_flowcell_undetermined, echo=FALSE}
undetermined_plot <- basic_plot(
                       flowcell_qc_info,
                       "undetermined_percentage",
                       "% Undetermined",
                       100,
                       "Non-Diagnostic",
                       qc_date_range,
                       "sequencer"
                     ) +
                     geom_hline(yintercept=8, linetype="dashed")
ggplotly(undetermined_plot)

basic_table(coefficients_annotation(flowcell_qc_info, "date", "undetermined_percentage", "sequencer"))
```
</div>
</div>


# Sample
### Q30
<div class = "row">
<div class = "column">
```{r diagnostic_sample_q30_sequencer, echo=FALSE}
q30_plot <- basic_plot(
              diagnostic_sample_qc_info,
              "sample_q30",
              "% Q30",
              100,
              "Diagnostic",
              qc_date_range,
              "sequencer"
            ) +
            geom_hline(yintercept=75, linetype="dashed")
ggplotly(q30_plot)

basic_table(coefficients_annotation(diagnostic_sample_qc_info, "date", "sample_q30", "sequencer"))
```
</div>
<div class = "column">
```{r nondiagnostic_sample_q30_sequencer, echo=FALSE}
q30_plot <- basic_plot(
              sample_qc_info,
              "sample_q30",
              "% Q30",
              100,
              "Non-Diagnostic",
              qc_date_range,
              "sequencer"
            ) +
            geom_hline(yintercept=75, linetype="dashed")
ggplotly(q30_plot)

basic_table(coefficients_annotation(sample_qc_info, "date", "sample_q30", "sequencer"))
```
</div>
</div>

<div class = "row">
<div class = "column">
```{r diagnostic_sample_q30_type, echo=FALSE}
q30_plot <- basic_plot(
              diagnostic_sample_qc_info,
              "sample_q30",
              "% Q30",
              100,
              "Diagnostic",
              qc_date_range,
              "type"
            ) +
            geom_hline(yintercept=75, linetype="dashed")
ggplotly(q30_plot)

basic_table(coefficients_annotation(diagnostic_sample_qc_info, "date", "sample_q30", "type"))
```
</div>
<div class = "column">
```{r nondiagnostic_sample_q30_type, echo=FALSE}
q30_plot <- basic_plot(
              sample_qc_info,
              "sample_q30",
              "% Q30",
              100,
              "Non-Diagnostic",
              qc_date_range,
              "type"
            ) +
            geom_hline(yintercept=75, linetype="dashed")
ggplotly(q30_plot)

basic_table(coefficients_annotation(sample_qc_info, "date", "sample_q30", "type"))
```
</div>
</div>

### Yield
<div class = "row">
<div class = "column">
```{r diagnostic_sample_yield_sequencer, echo=FALSE}
yield_plot <- basic_plot(
                diagnostic_sample_qc_info,
                "sample_yield",
                "GBase Yield",
                NA,
                "Diagnostic",
                qc_date_range,
                "sequencer"
              ) +
              geom_hline(yintercept=1, linetype="dashed")
ggplotly(yield_plot)

basic_table(coefficients_annotation(diagnostic_sample_qc_info, "date", "sample_yield", "sequencer"))
```
</div>
<div class = "column">
```{r nondiagnostic_sample_yield_sequencer, echo=FALSE}
yield_plot <- basic_plot(
                nondiagnostic_sample_qc_info,
                "sample_yield",
                "GBase Yield",
                NA,
                "Non-Diagnostic",
                qc_date_range,
                "sequencer"
              ) +
             geom_hline(yintercept=1, linetype="dashed")
ggplotly(yield_plot)

basic_table(coefficients_annotation(nondiagnostic_sample_qc_info, "date", "sample_yield", "sequencer"))
```
</div>
</div>

<div class = "row">
<div class = "column">
```{r diagnostic_sample_yield_type, echo=FALSE}
yield_plot <- basic_plot(
                diagnostic_sample_qc_info,
                "sample_yield",
                "GBase Yield",
                NA,
                "Diagnostic",
                qc_date_range,
                "type"
              )  +
              geom_hline(yintercept=1, linetype="dashed")
ggplotly(yield_plot)

basic_table(coefficients_annotation(diagnostic_sample_qc_info, "date", "sample_yield", "type"))

```
</div>
<div class = "column">
```{r nondiagnostic_sample_yield_type, echo=FALSE}
yield_plot <- basic_plot(
                nondiagnostic_sample_qc_info,
                "sample_yield",
                "GBase Yield",
                NA,
                "Non-Diagnostic",
                qc_date_range,
                "type"
              ) +
              geom_hline(yintercept=1, linetype="dashed")
ggplotly(yield_plot)

basic_table(coefficients_annotation(nondiagnostic_sample_qc_info, "date", "sample_yield", "type"))
```
</div>
</div>

# Run
<div class = "row">
<div class = "column">
```{r run_median_coverage, echo=FALSE}
cov_plot <- basic_plot(
              run_qc_info,
              "cov",
              "Coverage",
              NA,
              "Median Coverage",
              qc_date_range,
              "type"
            )
ggplotly(cov_plot)

basic_table(coefficients_annotation(run_qc_info, "date", "cov", "type"))
```
</div>
<div class = "column">
```{r run_tumor_bqr_rate, echo=FALSE}
bqr_plot <- basic_plot(
              run_qc_info,
              "bqr",
              "% BQR",
              NA,
              "BQR Error Rate",
              qc_date_range,
              "type"
            )
ggplotly(bqr_plot)

basic_table(coefficients_annotation(run_qc_info, "date", "bqr", "type"))
```
</div>
</div>

# Stats

Number of flowcells: `r nrow(flowcell_qc_info)`
Number of flowcells with diagnostic data: `r nrow(diagnostic_flowcell_qc_info)`

Number of samples: `r nrow(sample_qc_info)`
Number of diagnostic samples: `r nrow(diagnostic_sample_qc_info)`

### Numbers per Submission/Study (Diagnostics)
```{r stats_numbers, echo=FALSE}
submission_numbers <- diagnostic_sample_qc_info %>%
                      count(submission, type, sort = TRUE) %>%
                      pivot_wider(names_from = type, values_from = n) %>%
                      mutate(total=tumor+ref)
basic_table(submission_numbers)
```

### Distribution Flowcells
<div class = "row">
<div class = "column">
```{r stats_diagnostic_flowcells, echo=FALSE}
samples_plot <- ggplot(diagnostic_flowcell_qc_info, aes(date, fill=sequencer)) +
                geom_bar(stat="count") +
                labs(y="Count", x="Date/Time", fill="sequencer", title="Diagnostic") +
                scale_x_date(date_labels="%m-%Y") +
                xlim(as.Date(qc_date_range, format="%Y-%m-%d")) +
                theme_minimal() +
                theme(legend.position="right", axis.text.x=element_text(size=6))
ggplotly(samples_plot)
```
</div>
<div class = "column">
```{r stats_nondiagnostic_flowcells, echo=FALSE}
samples_plot <- ggplot(nondiagnostic_flowcell_qc_info, aes(date, fill=sequencer)) +
                geom_bar(stat="count") +
                labs(y="Count", x="Date/Time", fill="sequencer", title="Non-Diagnostic") +
                scale_x_date(date_labels="%m-%Y") +
                xlim(as.Date(qc_date_range, format="%Y-%m-%d")) +
                theme_minimal() +
                theme(legend.position="right", axis.text.x=element_text(size=6))
ggplotly(samples_plot)
```
</div>
</div>

### Distribution Diagnostic Samples
<div class = "row">
<div class = "column">
```{r stats_samples_sequencer, echo=FALSE}
samples_plot <- ggplot(diagnostic_sample_qc_info, aes(date, fill=sequencer)) +
                geom_bar(stat="count") +
                labs(y="Count", x="Date/Time", fill="sequencer", title="Diagnostic") +
                scale_x_date(date_labels="%m-%Y") +
                xlim(as.Date(qc_date_range, format="%Y-%m-%d")) +
                theme_minimal() +
                theme(legend.position="right", axis.text.x=element_text(size=6))
ggplotly(samples_plot)
```
</div>
<div class = "column">
```{r stats_samples_type, echo=FALSE}
samples_plot <- ggplot(diagnostic_sample_qc_info, aes(date, fill=type)) +
                geom_bar(stat="count") +
                labs(y="Count", x="Date/Time", fill="type", title="Diagnostic") +
                scale_x_date(date_labels="%m-%Y") +
                xlim(as.Date(qc_date_range, format="%Y-%m-%d")) +
                theme_minimal() +
                theme(legend.position="right", axis.text.x=element_text(size=6))
ggplotly(samples_plot)
```
</div>
</div>


### Lab Failures
<div class = "row">
<div class = "column">
```{r iso_fails, echo=FALSE}
samples_plot <- ggplot(iso_fail_info, aes(date)) +
                geom_bar(stat="count") +
                labs(y="Count", x="Date/Time", title="Isolation") +
                scale_x_date(date_labels="%m-%Y") +
                xlim(as.Date(qc_date_range, format="%Y-%m-%d")) +
                theme_minimal() +
                theme(legend.position="right", axis.text.x=element_text(size=6))
ggplotly(samples_plot)
```
</div>
<div class = "column">
```{r prep_fails, echo=FALSE}
samples_plot <- ggplot(prep_fail_info, aes(date)) +
                geom_bar(stat="count") +
                labs(y="Count", x="Date/Time", title="Prep") +
                scale_x_date(date_labels="%m-%Y") +
                xlim(as.Date(qc_date_range, format="%Y-%m-%d")) +
                theme_minimal() +
                theme(legend.position="right", axis.text.x=element_text(size=6))
ggplotly(samples_plot)
```
</div>
</div>


# GIAB
<div class = "row">
<div class = "column">
```{r giab_q30, echo=FALSE}
q30_plot <- basic_plot(
              giab_info,
              "Q30",
              "% Q30",
              100,
              "Q30",
              giab_date_range,
              "Machine"
            )
ggplotly(q30_plot)

basic_table(coefficients_annotation(giab_info, "date", "Q30", "Machine"))
```
</div>
<div class = "column">
```{r giab_mapped, echo=FALSE}
snv_plot <- basic_plot(
              giab_info,
              "Mapped",
              "Mapped %",
              100,
              "Mapped",
              giab_date_range,
              "Machine"
            )
ggplotly(snv_plot)

basic_table(coefficients_annotation(giab_info, "date", "Mapped", "Machine"))
```
</div>
</div>

<div class = "row">
<div class = "column">
```{r giab_excluded, echo=FALSE}
excluded_plot <- basic_plot(
                   giab_info,
                   "Excluded",
                   "Total Excluded %",
                   100,
                   "Total Excluded",
                   giab_date_range,
                   "Machine"
                 )
ggplotly(excluded_plot)

basic_table(coefficients_annotation(giab_info, "date", "Excluded", "Machine"))
```
</div>
<div class = "column">
```{r giab_duplicates, echo=FALSE}
snv_plot <- basic_plot(
              giab_info,
              "Duplicates",
              "Duplicates Excluded Duplicates %",
              100,
              "Duplicates Excluded",
              giab_date_range,
              "Machine"
            )
ggplotly(snv_plot)

basic_table(coefficients_annotation(giab_info, "date", "Duplicates", "Machine"))
```
</div>
</div>

# Data
### GIAB
GIAB data can be found in `HMF-FOR-096 SEQUENCING SYSTEMS VERIFICATIONS v2.0.xlsx`.

### Flowcell QC Info
Flowcell QC info data is an aggregation of `Sample QC Info` on the column `flowcell`.

### Sample QC Info
```{r data_sample, echo=FALSE}
datatable(sample_qc_info,
          selection="multiple",
          escape=FALSE,
          extensions='Buttons',
          rownames=TRUE,
          options = list(dom='Bfrtip', buttons=c('csv', 'excel'), pageLength=15, lengthChange=FALSE)
          )
```
### Run QC Info
```{r data_run, echo=FALSE}
datatable(run_qc_info,
          selection="multiple",
          escape=FALSE,
          extensions='Buttons',
          rownames=TRUE,
          options = list(dom='Bfrtip', buttons=c('csv', 'excel'), pageLength=15, lengthChange=FALSE)
)
```