---
title: "GRIDSS clustering"
output:
  html_document:
    df_print: paged
---

```{r}
#library(purple)
library(RMySQL)
library(tidyverse)
library(Biostrings)
library(StructuralVariantAnnotation)
library(testthat)
tmpwd=getwd()
setwd("../gridss/")
source("libgridss.R")
setwd(tmpwd)
remove(tmpwd)
source("libSvAnalyser.R")
```

Load clustering data from file and GRIDSS calls from DB
```{r}
svData = read.csv('../../cluster/CLUSTER_GRIDSS_v2.csv') %>%
  mutate(source="proximity")
```

Load GRIDSS calls from DB

```{r}
dbgridss = dbConnect(MySQL(), dbname='gridss_test')
cndf = query_all_copy_numer(dbgridss)
cngr = to_cn_gr(cndf)
svdf = query_somatic_structuralVariants(dbgridss)
svgr = to_sv_gr(svdf)
svgr$cnid = annotate_sv_with_cnv_id(cngr, svgr, maxgap=1000)
if (any(is.na(svgr$cnid))) {
  bad_svid = svgr$id[is.na(svgr$cnid)]
  warning(paste("Missing matching segment boundary for", length(bad_svid), "SVs. Removing from analysis. TODO: work out what happened to these bounds"))
  svdf = svdf %>% filter(!(id %in% bad_svid))
  svData = svData %>% filter(!(Id %in% bad_svid))
  svgr = svgr[!(svgr$id %in% bad_svid)]
}
```


Compare clustering to GRIDSS linkages
```{r}
gridss_links = gridss_sv_links(svdf)
svdf$gridss_cluster = cluster_from_links(svdf, gridss_links)
svdf = svdf %>% group_by(gridss_cluster) %>%
  mutate(source="gridss") %>%
  mutate(ClusterCount=n()) %>%
  ungroup()

gridss_clustered_sv_pairs = svdf %>% dplyr::select(sampleId, gridss_cluster, id) %>%
  inner_join(
    svdf %>% dplyr::select(sampleId, gridss_cluster, id),
    by=c("sampleId"="sampleId", "gridss_cluster"="gridss_cluster"),
    suffix=c("", "2")) %>%
  filter(id != id2)
proximity_clustered_sv_pairs = svData %>% dplyr::select(SampleId, ClusterId, Id) %>%
  inner_join(
    svData %>% dplyr::select(SampleId, ClusterId, Id),
    by=c("SampleId"="SampleId", "ClusterId"="ClusterId"),
    suffix=c("", "2")) %>%
  filter(Id != Id2)

gridss_unique_clustered_sv_pairs = anti_join(gridss_clustered_sv_pairs, proximity_clustered_sv_pairs,
  by=c("id"="Id", "id2"="Id2"))

svdf %>% dplyr::select(SampleId=sampleId, ClusterCount, source) %>%
  bind_rows(svData %>% dplyr::select(SampleId, ClusterCount, source)) %>%
ggplot() +
  aes(x=ClusterCount) +
  geom_histogram() +
  facet_wrap(~ source) +
  scale_y_log10() +
  scale_x_log10() +
  labs(title="Size of cluster", x="cluster size", y="SV count")

svData = svData %>% mutate(missing_gridss_link=paste0(SampleId,"-", ClusterId) %in% (svData %>%
  group_by(SampleId, ClusterId) %>%
  summarise(has_missing = any(Id %in% gridss_unique_clustered_sv_pairs$id)) %>%
  filter(has_missing) %>%
  mutate(temp=paste0(SampleId, "-", ClusterId)) %>%
  pull(temp)))

data.frame(
  gridss_unique_clustered_breakpoints=svdf %>% filter(!is.na(endChromosome)) %>% filter(id %in% gridss_unique_clustered_sv_pairs$id) %>% nrow(),
  gridss_unique_clustered_breakends=svdf %>% filter(is.na(endChromosome)) %>% filter(id %in% gridss_unique_clustered_sv_pairs$id) %>% nrow(),
  gridss_clusters=svdf %>% group_by(gridss_cluster) %>% summarise() %>% nrow(),
  proxmity_clusters=svData %>% group_by(SampleId, ClusterId) %>% summarise() %>% nrow())


```



Visualise sample
```{r}
library(visNetwork)
sample = "CPCT02010255T"
segmentSupportShape = function(support) {
  ifelse(support %in% c("TELOMERE", "CENTROMERE"), "triangle",
    ifelse(support == "BND", "square",
      ifelse(support == "MULTIPLE", "circle",
        "star")))
}
nodes = bind_rows(
  # start
  cndf %>% 
    filter(sampleId == sample) %>%
    mutate(
      label=round(copyNumber, 1),
      size=pmax(1, copyNumber),
      id=paste0(id, "-"),
      shape=segmentSupportShape(segmentStartSupport)),
  # end
  cndf %>%
    filter(sampleId == sample) %>%
    mutate(
      size=copyNumber,
      id=paste0(id, "+"),
      shape=segmentSupportShape(segmentEndSupport))
  ) %>% mutate(
    color="lightblue")
edges = bind_rows(
  # internal segment edges
  cndf %>%
    filter(sampleId == sample) %>%
    mutate(
      from=paste0(paste0(id, "-")),
      to=paste0(paste0(id, "+")),
      color="lightblue",
      width=copyNumber,
      length=log10(end - start) + 1,
      title=paste0(chromosome, ":", start, "-", end, " (", end - start, "bp)"),
      smooth=FALSE
      ) %>%
    dplyr::select(from, to, color, width, length, title, smooth),
  # Reference edges
  cndf %>%
    filter(sampleId == sample) %>%
    group_by(sampleId, chromosome) %>%
    arrange(start) %>%
    mutate(nextid=lead(id)) %>%
    ungroup() %>%
    filter(!is.na(nextid)) %>%
    mutate(color=ifelse(segmentEndSupport == "CENTROMERE", "lightgreen", "green"),
      from=paste0(paste0(id, "+")),
      to=paste0(paste0(nextid, "-")),
      label=NA,
      width=2,
      length=NA,
      title=NA,
      smooth=FALSE) %>%
    dplyr::select(from, to, color, label, width, length, title, smooth),
  #
  svgr %>% as.data.frame() %>%
    filter(sampleId == sample) %>%
    group_by(sampleId, id) %>%
    arrange(seqnames, start) %>%
    mutate(
      partner_orientation=lead(strand),
      partner_cnid=lead(cnid)) %>%
    ungroup() %>%
    filter(!is.na(partner_cnid)) %>%
    mutate(
      color="red",
      from=paste0(paste0(cnid, strand)),
      to=paste0(paste0(partner_cnid, partner_orientation)),
      label=round(ploidy, 1),
      width=ploidy,
      length=NA,
      title=NA,
      smooth=TRUE) %>%
    dplyr::select(from, to, color, label, width, length, title, smooth))

rescaling = list(width=5, length=3, size=1)
visNetwork(
    nodes %>% mutate(size=size * rescaling$size),
    edges %>% mutate(width=width * rescaling$width, length=length * rescaling$length),
    height = "1000px", width = "100%") %>%
  visLayout(randomSeed = 12)

``




